---
title: "Reproducible Research: Peer Assessment 1"
author: "amit"
date: "6/27/2020"
output: html_document
---
## mean total number of steps taken per day
LOading the data
```{r,echo=TRUE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,message=FALSE, warning=FALSE)
library(tidyr)
library(dplyr)
library(data.table)
library(ggplot2)
library(ggrepel)
library(lubridate)

activity <- read.csv("activity.csv")
```
Calculate the total number of steps taken per day
```{r}
activity$date <- ymd(as.character(activity$date))
activity$HMS <- hms(paste(activity$interval%/%100,":",activity$interval-(activity$interval%/%100)*100,":",0))
step_by_day <- aggregate(steps~date  ,data= activity,FUN="sum")
knitr::kable(summary(step_by_day))
```
histogram of the total number of steps taken each day
```{r}
ggplot(step_by_day,aes(date))+geom_histogram(aes(steps),bins = 30)
```
mean and median of the total number of steps taken per day
```{r}
mean(step_by_day$steps)
median(step_by_day$steps)
```
##average daily activity pattern
time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
average_step_by_5mn <- aggregate(steps ~ interval ,data= activity,FUN="mean")
ggplot(average_step_by_5mn,aes(x = interval/100,y= steps))+geom_line()+
  labs (title= "Average number of step by 5mn interval", x="5mn interval (in hour)")
```

5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
max_average_number_of_steps <- average_step_by_5mn[which(average_step_by_5mn$steps == max(average_step_by_5mn$steps, na.rm = TRUE)), ]
max_average_number_of_steps
```
##max_average_number_of_steps
 total number of missing values in the dataset

```{r}
total_rows_with_NAs <-  nrow(activity)-sum(complete.cases(activity))
```
All missing values in the dataset will be replaced by the mean of the corresponding 5-minutes interval
```{r}
activity$modify<- is.na(activity$steps)
activity <-merge(activity,average_step_by_5mn,by = "interval" )
activity$steps.x[activity$modify] <- activity$steps.y[activity$modify]
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_new<- data.frame(interval=activity$interval,steps = activity$steps.x,date=as.Date(activity$date))
knitr::kable(summary(activity_new))
```

```{r}
knitr::kable(head(activity_new))
```
 histogram of the total number of steps taken each day

```{r}
total_step_by_day <- aggregate(steps ~ date ,data= activity_new,FUN="sum")
ggplot(total_step_by_day,aes(date))+geom_histogram(aes(steps),bins = 30)
```

```{r}
mean_before_correction <- mean(step_by_day$steps)
mean_after_correction <-mean(total_step_by_day$steps)
median_before_correction <- median(step_by_day$steps)
median_after_correction <- median(total_step_by_day$steps)
```
##differences in activity patterns between weekdays and weekends
new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
week_days <- c("lundi","mardi","mercredi","jeudi","vendredi")
weekend_days <- c("samedi","dimanche")
day_type <- weekdays(activity_new$date) %in% week_days
activity_new$day_type[day_type]= "weekday" 
activity_new$day_type[!(day_type)]= "weekend"
activity_new$day_type <- as.factor(activity_new$day_type)
knitr::kable(head(activity_new))
```
panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
average_step_by_5mn_daytype <- aggregate(steps ~ interval + day_type ,data= activity_new,FUN=mean)
ggplot(data = average_step_by_5mn_daytype, aes(x= interval/100, y = steps))+geom_line()+facet_grid(day_type~.)+
  labs (title= "Average number of step by interval and day type", x="5mn interval (in hour)")
```
